### 2020-04-01 Thesis-Chapter2-Background.tex
Finish section of first crypto war

Despite the failure of the Clipper Chip, the debate over access to strong encryption continued, primarily focused on
various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
it was S.1726, the Pro-C0DE Act of 1996, to abolish the export controls on encryption software \cite{burns_s1726_1996}.
In the House, it was HR.3011, the SAFE Act of 1996, to explicitly allow arbitrarily strong encryption for all legal
activity \cite{goodlatte_hr3011_1996}. Indeed, the SAFE Act went further, including ``provisions that would have
barred the government from creating a mandatory key escrow system and would also have removed export restrictions on
most generally available software and hardware containing encryption'' \cite{kehl_right_2015}. Zimmermann, the author of
PGP, who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. 
% TODO: "The the Pro-C0DE Act passed, but the SAFE Act did not." -- WRONG! Neither passed, this was also achieved by executive order!

@@

Two developments sewed the seeds for the Second Crypto War. The first was actually a non-development: when the Clinton
administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. This meant
that rather than being written in the pen of law, encryption policy was written in the pencil of executive order. The
second development was the events of and reaction to the morning of September 11, 2001. (resources on 9/11:
\cite{shamsi_2011}, \cite{bloss_escalating_2007}.)

@@

\begin{itemize}
    \item The All Writs Act \cite{congress_1789}
    \item The Bill of Rights: Amendments I, IV, and V \cite{madison_1791}
    \item The Arms Export Control Act (AECA) of 1976 established export controls over strong encryption techniques
        introduced ITAR (International Traffic in Arms Regulations) \cite{morgan_hr13680_1976} \cite{kehl_right_2015}
    \item Foreign Intelligence Surveillance (FISA) Act \cite{rodino_1978}
    \item The Computer Fraud and Abuse Act (CFAA) of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
    \item Electronic Communications Privacy Act (ECPA) of 1986 \cite{kastenmeier_hr4952_1986}
    \item The Communications Assistance for Law Enforcement Act (CALEA) of 1994 \cite{edwards_hr4922_1994}
    \item Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism
        (USA PATRIOT ACT) Act of 2001 \cite{sensenbrenner_2001}
    \item FISA Amendments Act of 2008 \cite{reyes_hr6304_2008}
    \item Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring
        (USA FREEDOM) Act of 2015 \cite{sensenbrenner_2015}
    \item \dots
\end{itemize}

>>>>

Despite the failure of the Clipper Chip, the debate over access to strong encryption continued, primarily focused on
various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
it was S.1726, the Pro-C0DE Act of 1996, to abolish the export controls on encryption software \cite{burns_s1726_1996}.
In the House, it was HR.3011, the SAFE Act of 1996, to explicitly allow arbitrarily strong encryption for all legal
activity \cite{goodlatte_hr3011_1996}. Indeed, the SAFE Act went further, including ``provisions that would have
barred the government from creating a mandatory key escrow system and would also have removed export restrictions on
most generally available software and hardware containing encryption'' \cite{kehl_right_2015}. Zimmermann, the author of
PGP, who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. In November of that year, the Clinton administration
released an executive order effectively removing export controls on strong encryption products, along the lines of the
Pro-C0DE Act \cite{clinton_1996}.

@@

Two developments sewed the seeds for the Second Crypto War. The first was actually a non-development: when the Clinton
administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. Also un-passed
were the related Pro-C0DE Act, Encrypted Communications Privacy Act (not to be confused with the \textit{Electronic}
Communications Privacy Act (ECPA) of 1986), and Encryption Protects the Rights of Individuals from Violation and Abuse
in CYberspace (E-PRIVACY) Act \cite{leahy_s376_1997} \cite{ashcroft_s2067_1998}. This legislative failure meant that
rather than being written in the pen of law, encryption policy was written in the pencil of executive order.

The second development was the events of and reaction to the morning of September 11, 2001. (resources on 9/11:
\cite{shamsi_2011}, \cite{bloss_escalating_2007}.)

@@

\begin{itemize}
    \item The All Writs Act \cite{congress_1789}
    \item The Bill of Rights: Amendments I, IV, and V \cite{madison_1791}
    \item The Arms Export Control Act (AECA) of 1976 established export controls over strong encryption techniques
        introduced ITAR (International Traffic in Arms Regulations) \cite{morgan_hr13680_1976} \cite{kehl_right_2015}
    \item Foreign Intelligence Surveillance (FISA) Act \cite{rodino_1978}
        (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
        % (the FBI almost lost some provisions; follow up on https://apnews.com/110ff653e98478da66765897dde3d613,
        %     https://www.washingtontimes.com/news/2020/mar/15/adam-schiff-fisa-reform-intervention-protects-fbi-/,
        %     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)
    \item The Computer Fraud and Abuse Act (CFAA) of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
    \item Electronic Communications Privacy Act (ECPA) of 1986 \cite{kastenmeier_hr4952_1986}
    \item The Communications Assistance for Law Enforcement Act (CALEA) of 1994 \cite{edwards_hr4922_1994}
    \item Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism
        (USA PATRIOT ACT) Act of 2001 \cite{sensenbrenner_2001}
    \item FISA Amendments Act of 2008 \cite{reyes_hr6304_2008}
    \item Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring
        (USA FREEDOM) Act of 2015 \cite{sensenbrenner_2015}
    \item \dots
\end{itemize}
